---
title: "TDOR_data_explore"
author: "Paul Brennan"
date: "12/11/2018"
output: html_document
---

## Getting the data in...
So I want to explore the TDOR data and try to make some visualisations.
First step getting the data into R...
```{r downloaded_file, eval=FALSE}
# pull in a downloaded data file...
data <- read.csv("../tdor_2018_01.csv")
library(tidyverse)
data2<- read_csv("../tdor_2018_01.csv")
# probably slightly better - fewer factors!
```


Can I get it straight from bitbucket?
```{r getting_data, eval=FALSE}
# can I pull directly from bitbucket?
url <- "https://bitbucket.org/annajayne/tdor_data/raw/df40d56688757b0ab388c79b8c15dd6512e3a1d7/Data/TDoR%202018/tdor_2018_01.csv"

data3 <- read_csv(url)
```

Yes, this works but the url is a bit of a challenge and doesn't seem to be
able to be made programmatically...

Does a shorter url work?
```{r shorter_bitbucket_url, eval=FALSE}
# can I use a shorter url?
shorter_url <- "https://bitbucket.org/annajayne/tdor_data/raw/Data/TDoR%202018/tdor_2018_01.csv"
data4 <- read_csv(shorter_url)
```

NO!

## Full download to computer and local explore
Download all the data from [bitbucket page](https://bitbucket.org/annajayne/tdor_data/downloads/).
The actual url for direct download is [here](https://bitbucket.org/annajayne/tdor_data/get/ff2f91a187c2.zip)
Downloads a zip file with all the data including pictures - quite large. 

I have done this but it could probably be done straight into R.

Well, I will explore the 2018 data:
```{r read_local_files, eval=FALSE}
setwd("/Users/paulbrennan/Documents/TDOR_data_et_al/annajayne-tdor_data-ff2f91a187c2/Data/TDoR 2018")

file_names <- list.files()
file_names <- file_names[2:13]

data_2018 <- lapply(file_names, read_csv)

```
This produces a Large list with 12 elements.

Can I combine these into one data frame?
```{r combine_list, eval = FALSE}
# throws error due to age...
data_2018_df <- bind_rows(data_2018)
```

Age of one person is a range which means that this row of data is made into a 
character and cannot combine all into one tibble. 
Alter age of one person
```{r change_age, eval = FALSE}
# throws error due to age...
data_2018[[5]][13,2] <- 32
data_2018[[5]]$Age <- as.integer(data_2018[[5]]$Age)
data_2018_df <- bind_rows(data_2018)
```

Still throws an error due to similar issue in data_2018[7] and data_2018[9]
```{r explore_age_data, eval=FALSE}
data_2018[7]
# Age is character
data_2018[[7]]$Age
# 25th row is 40-50
# as this is a generic issue, we need to decide on a solution...
```

Possible solution:
during import turn Age into two columns with age_lower and age_upper. 
Following advice from [R for Data Science](https://r4ds.had.co.nz/tidy-data.html)
```{r separate_age_and_bind_rows, eval = FALSE}

data_2018[[5]]<- separate(data_2018[[5]], Age, into = c("Age", "Age_high"))
data_2018[[5]]$Age <- as.numeric(data_2018[[5]]$Age)
data_2018[[7]]<- separate(data_2018[[7]], Age, into = c("Age", "Age_high"))
data_2018[[7]]$Age <- as.numeric(data_2018[[7]]$Age)
data_2018[[9]]<- separate(data_2018[[9]], Age, into = c("Age", "Age_high"))
data_2018[[9]]$Age <- as.numeric(data_2018[[9]]$Age)
data_2018_df <- bind_rows(data_2018)
# there has to be a better way to do this rather than hard coding!

```

Next step to explore data_2018_df
```{r eliminate_empty_rows, eval = FALSE}
str(data_2018_df)
# get rid of empty rows
data_2018_nagone <- filter(data_2018_df, Name != "NA")

```
Checking this data reveals a challenge with the March 2018 data. 
Different name of column - date of death rather than just date and 
date column seems to have 
```{r explore_march_data, eval = FALSE}
colnames(data_2018[[6]])
# change column names
names <- colnames(data_2018[[6]])
names

colnames(data_2018[[5]])
names[5] <- "Date"

data_2018[[6]]$Date
# it's really Cause of death

names[11] <- "Cause of death"
names
colnames(data_2018[[6]]) <- names
data_2018_df <- bind_rows(data_2018)

# get rid of empty rows
data_2018_nagone <- filter(data_2018_df, Name != "NA")

# get rid of empty column at end...
data_2018_nagone$X14 <- NULL
data <- data_2018_nagone
```

OK so this looks correct. 353 obs of 14 variables. 

Make date a date using lubridate package. 
```{r dates, eval=FALSE}
library(lubridate)
data$Date <- dmy(data$Date)
```

After this munging...
Save the data...

```{r save_data, eval=FALSE}
write_csv(data, "TDOR_for_R_20181113")
```

I've put this on Github to facilitate a little visualisation
```{r read_in_data}

data <- readr::read_csv("https://raw.githubusercontent.com/brennanpincardiff/tg_data_working/master/TDOR_for_R_20181113.csv")
```


## Try plotting some data
Deaths across the year
```{r deaths_over_year}
library(ggplot2)
ggplot(data, aes(Date)) + geom_bar() +
    ggtitle("Deaths across the year")
```

Deaths by age
```{r deaths_by_age}
ggplot(data, aes(Age)) + geom_bar() +
    ggtitle("Deaths by age")
```



Deaths by country
```{r deaths_by_country}
ggplot2::ggplot(data, aes(Country)) + geom_bar() + 
    ggtitle("Deaths by country") +
    theme(axis.text.x = element_text(angle=45, hjust=1))
```